Seismic Time Series Analysis\\Statistics for Spatio-Temporal Data\\\textit{-- project draft --}

### 2020-06-09 project.after.Rmd
use original model for fitting lockdown data

fit.name <- paste0("fit", paste0(chosen.model, collapse = ""))
forecast.name <- paste0("forecast", paste0(chosen.model, collapse = ""))
cat("SARIMA(", paste0(chosen.model[1:3], collapse = ","), ")(", paste0(chosen.model[4:6], collapse = ","), ")[", frequency(final.training), "]", sep = "")
if (!exists("final.model")) {
  cat(" fitting...")
  final.model <- Arima(final.training, chosen.model[1:3], chosen.model[4:6], lambda=chosen.lambda)
  cat("DONE")
}
if (!exists("final.forecast")) {
  cat(" forecasting...")
  final.forecast <- forecast(final.model, h=168*8, lambda=chosen.lambda)
  cat("DONE")
}
cat("\n")
plotting.model <- final.model
plotting.forecast <- final.forecast
print(plotting.model)
print(autoplot(plotting.model) + labs(subtitle = plotting.forecast$method))
checkresiduals(plotting.model)
residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*4)) + scale_x_continuous() + labs(title = paste0("Residuals of ", plotting.forecast$method))
residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*4)) + scale_x_continuous() + labs(title = paste0("Residuals of ", plotting.forecast$method))
print(plot_grid(residuals.acf, residuals.pacf, nrow=2))
print(plot_grid(
  autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=24)) + scale_x_continuous() + labs(title = paste0("Residuals of ", plotting.forecast$method)),
  residuals.acf,
  autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=24)) + scale_x_continuous() + labs(title = paste0("Residuals of ", plotting.forecast$method)),
  residuals.pacf,
  nrow = 2
))
print(plot_forecast(plotting.forecast) + autolayer(final.testing, series="Original") + ylab("Mean hourly movement (nm)") + xlab("Weeks") + geom_vline(aes(xintercept=as.integer(as_datetime('2020-03-14T00:00:00') - seis.train.start)/7+1), linetype=2))
plots <- show_overlap_and_diff(
  final.training, plotting.forecast$fitted, xlab=seasonality.xlab,
  base.desc = "training set",
  compare.desc = paste0("refit with ", plotting.forecast$method),
  overlap.ylab = "Mean hourly movement (nm)",
  diff.title = "Difference between ts",
  diff.ylab = "nm", diff.ylab.sec="%",
  diff.ylab.sec.accuracy = .01
)
print(plot_grid(plots[[1]], plots[[2]], nrow = 2))

cat("Test set inside the CI\n")
cat("80% CI \t ", print_dec(mean(plotting.forecast$lower[,1] < final.testing & final.testing < plotting.forecast$upper[,1], na.rm = T)*100), "%\n")
cat("95% CI \t ", print_dec(mean(plotting.forecast$lower[,2] < final.testing & final.testing < plotting.forecast$upper[,2], na.rm = T)*100), "%\n")

plots <- show_overlap_and_diff(
  final.testing, plotting.forecast$mean, xlab=seasonality.xlab,
  base.desc = "test set",
  compare.desc = paste0("forecast with ", plotting.forecast$method),
  overlap.ylab = "Mean hourly movement (nm)",
  diff.title = "Difference between ts",
  diff.ylab = "nm", diff.ylab.sec="%",
  diff.ylab.sec.accuracy = .01
)
print(plot_grid(plots[[1]] + geom_vline(aes(xintercept=as.integer(as_datetime('2020-03-14T00:00:00') - seis.train.start)/7+1), linetype=2),
                plots[[2]] + geom_vline(aes(xintercept=as.integer(as_datetime('2020-03-14T00:00:00') - seis.train.start)/7+1), linetype=2),
                nrow = 2
                ))

print(accuracy(plotting.forecast, final.testing))
# qqPlot(plotting.model$residuals, main = paste0("Q-Q plot for ", plotting.forecast$method), xlab = "residuals")
print(gg_qqplot(plotting.model$residuals) + 
    labs(title=paste0("Q-Q plot for ", plotting.forecast$method, " residuals")) + ylab("Residuals quantiles")
)

>>>>

fit.name <- paste0("fit", paste0(chosen.model, collapse = ""))
forecast.name <- paste0("forecast", paste0(chosen.model, collapse = ""))
cat("SARIMA(", paste0(chosen.model[1:3], collapse = ","), ")(", paste0(chosen.model[4:6], collapse = ","), ")[", frequency(final.training), "]", sep = "")
if (!exists("full.forecast")) {
  cat(" forecasting...")
  full.forecast <- forecast(models[[fit.name]], h=168*26, lambda = chosen.lambda)
  cat("DONE")
}
cat("\n")
plotting.model <- models[[fit.name]]
plotting.forecast <- full.forecast
print(plotting.model)
print(autoplot(plotting.model) + labs(subtitle = plotting.forecast$method))
checkresiduals(plotting.model)
residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*4)) + scale_x_continuous() + labs(title = paste0("Residuals of ", plotting.forecast$method))
residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*4)) + scale_x_continuous() + labs(title = paste0("Residuals of ", plotting.forecast$method))
print(plot_grid(residuals.acf, residuals.pacf, nrow=2))
print(plot_grid(
  autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=24)) + scale_x_continuous() + labs(title = paste0("Residuals of ", plotting.forecast$method)),
  residuals.acf,
  autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=24)) + scale_x_continuous() + labs(title = paste0("Residuals of ", plotting.forecast$method)),
  residuals.pacf,
  nrow = 2
))
print(plot_forecast(plotting.forecast) + autolayer(ts_c(`Before March '20`=seis.h.test, `From March '20`=final.testing), series="Original") + ylab("Mean hourly movement (nm)") + xlab("Weeks") + geom_vline(aes(xintercept=as.integer(as_datetime('2020-03-14T00:00:00') - seis.train.start)/7+1), linetype=2))
plots <- show_overlap_and_diff(
  seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
  base.desc = "training set",
  compare.desc = paste0("refit with ", plotting.forecast$method),
  overlap.ylab = "Mean hourly movement (nm)",
  diff.title = "Difference between ts",
  diff.ylab = "nm", diff.ylab.sec="%",
  diff.ylab.sec.accuracy = .01
)
print(plot_grid(plots[[1]], plots[[2]], nrow = 2))

cat("Test set inside the CI\n")
cat("CI   ", "Overall ", "Before  ", "After  \n")
cat("80 % ",
    print_dec(mean(plotting.forecast$lower[,1] < c(seis.h.test, final.testing) & c(seis.h.test, final.testing) < plotting.forecast$upper[,1], na.rm = T)*100),
    "% ",
    print_dec(mean(plotting.forecast$lower[1:length(seis.h.test),1] < seis.h.test & seis.h.test < plotting.forecast$upper[1:length(seis.h.test),1], na.rm = T)*100),
    "% ",
    print_dec(mean(plotting.forecast$lower[(length(seis.h.test)+1):dim(plotting.forecast$lower)[1],1] < final.testing & final.testing < plotting.forecast$upper[(length(seis.h.test)+1):dim(plotting.forecast$lower)[1],1], na.rm = T)*100),
    "%\n")
cat("95 % ",
    print_dec(mean(plotting.forecast$lower[,2] < c(seis.h.test, final.testing) & c(seis.h.test, final.testing) < plotting.forecast$upper[,2], na.rm = T)*100),
    "% ",
    print_dec(mean(plotting.forecast$lower[1:length(seis.h.test),2] < seis.h.test & seis.h.test < plotting.forecast$upper[1:length(seis.h.test),2], na.rm = T)*100),
    "% ",
    print_dec(mean(plotting.forecast$lower[(length(seis.h.test)+1):dim(plotting.forecast$lower)[1],2] < final.testing & final.testing < plotting.forecast$upper[(length(seis.h.test)+1):dim(plotting.forecast$lower)[1],2], na.rm = T)*100),
    "%\n")

plots <- show_overlap_and_diff(
  c(seis.h.test, final.testing), plotting.forecast$mean, xlab=seasonality.xlab,
  base.desc = "test set",
  compare.desc = paste0("forecast with ", plotting.forecast$method),
  overlap.ylab = "Mean hourly movement (nm)",
  diff.title = "Difference between ts",
  diff.ylab = "nm", diff.ylab.sec="%",
  diff.ylab.sec.accuracy = .01
)
print(plot_grid(plots[[1]] + geom_vline(aes(xintercept=as.integer(as_datetime('2020-03-14T00:00:00') - seis.train.start)/7+1), linetype=2),
                plots[[2]] + geom_vline(aes(xintercept=as.integer(as_datetime('2020-03-14T00:00:00') - seis.train.start)/7+1), linetype=2),
                nrow = 2
                ))

plots <- show_overlap_and_diff(
  seis.h.test, window(plotting.forecast$mean, start=c(53,1), end=c(70,168)), xlab=seasonality.xlab,
  base.desc = "test set before lockdown",
  compare.desc = paste0("forecast with ", plotting.forecast$method),
  overlap.ylab = "Mean hourly movement (nm)",
  diff.title = "Difference between ts",
  diff.ylab = "nm", diff.ylab.sec="%",
  diff.ylab.sec.accuracy = .01
)
print(plot_grid(plots[[1]],
                plots[[2]],
                nrow = 2
                ))

plots <- show_overlap_and_diff(
  final.testing, window(plotting.forecast$mean, start=c(71,1), end=c(78,168)), xlab=seasonality.xlab,
  base.desc = "test set during lockdown",
  compare.desc = paste0("forecast with ", plotting.forecast$method),
  overlap.ylab = "Mean hourly movement (nm)",
  diff.title = "Difference between ts",
  diff.ylab = "nm", diff.ylab.sec="%",
  diff.ylab.sec.accuracy = .01
)
print(plot_grid(plots[[1]] + geom_vline(aes(xintercept=as.integer(as_datetime('2020-03-14T00:00:00') - seis.train.start)/7+1), linetype=2),
                plots[[2]] + geom_vline(aes(xintercept=as.integer(as_datetime('2020-03-14T00:00:00') - seis.train.start)/7+1), linetype=2),
                nrow = 2
                ))

print(accuracy(plotting.forecast, c(seis.h.test, final.testing)))
# qqPlot(plotting.model$residuals, main = paste0("Q-Q plot for ", plotting.forecast$method), xlab = "residuals")
print(gg_qqplot(plotting.model$residuals) + 
    labs(title=paste0("Q-Q plot for ", plotting.forecast$method, " residuals")) + ylab("Residuals quantiles")
)
